Experience Report: Haskell in Computational Biology

### 2012-03-11 mrfy_experience_report.tex
holes filled, pushing back

When performance has mattered, the members of our group, like other
computational biologists, have used~C++.
We~compare our experience with three tools:
\begin{itemize}
\item
Matt~\cite{Menke:2008wu} is used to create alignments like that shown
in \figref{alignment}.
It~comprises about 12,000 lines of~C++.
The only external tool or library it uses is \texttt{zlib}.
Its initial development took \emph{[HOW LONG?]}.
\item
SMURF
\cite{Menke:2010ti} is used \emph{[FOR WHAT?]}
It~comprises about \emph{[HOW MANY?]} lines of~C++.
Its initial development took \emph{[HOW LONG?]}.
\item
MRFy is used to detect homologous proteins in the presence of paired
beta strands.
It~comprises about \emph{[HOW MANY?]} lines of~Haskell
Its initial development took about three months of part-time work.
\remark{Loss of parallel structure---do we distinguish part time from
full time?}
\end{itemize}

It uses such
data structures such as mutable oct-trees, vectors, and arrays.
It~uses clever pointer arithmetic.
It relies on no external libraries beyond \texttt{zlib},
but contains 1,000 lines of protein-structure code that ought
to be in a library.

>>>>

When performance has mattered, the members of our group, like other
computational biologists, have used~C++.
We~compare our experience with three tools:
\begin{itemize}
\item
Matt~\cite{Menke:2008wu} is used to create alignments like that shown
in \figref{alignment}.
It~comprises about 12,000 lines of~C++.
The only external tool or library it uses is \texttt{zlib}.
Its initial development took \emph{two years}.
\item
SMURF
\cite{Menke:2010ti} is used to detect homologous proteins in the presence
of paired beta strands.
It~comprises about 9,000 lines of~C++.
Its initial development took a year and a half.
\item
MRFy is used to detect homologous proteins in the presence of paired
beta strands.
It~comprises about 1,200 lines of~Haskell
Its initial development took about three months.
\remark{Others were mostly full-time work, but no need to distinguish.}
\end{itemize}

Matt uses such
data structures such as mutable oct-trees, vectors, and arrays.
It~uses clever pointer arithmetic.
It relies on no external libraries beyond \texttt{zlib},
but contains 1,000 lines of protein-structure code that ought
to be in a library.
